---
title: "Simulation of soil temperature, vapour pressure and humidity at Newhaven"
author: "Michael Kearney"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Simulation of soil temperature and lizard activity at Kinchega NP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
 eval = TRUE, tidy.opts=list(width.cutoff=60), tidy=TRUE  
)
```

## Overview

This document compares observed and predicted soil temperature, vapour pressure and relative humidity to that measured by Danae Moore and Joe Schofield at Newhaven for the Great Desert Skink Project.

```{r}
library(NicheMapR)
library(dplyr)
```

## Run the model with daily gridded data for Australia

Run the model using historical daily gridded data for Australia from 2013 to 2014.

First set up some soil properties for the site.

```{r}
#source("c:/git/micro_australia/get.soil.R")
ystart <- 2013# start year
yfinish <- 2014# end year
nyears<-yfinish-ystart+1# integer, number of years for which to run the microclimate model
loc <- c(131.183806, -22.797235)
longlat <- rbind(loc)
```

Now set the soil properties. First retrieve soil properties from Soil and Landscape Grid of Australia (SLGA)

```{r, eval=FALSE}
prevdir<-getwd()
setwd('x:')
cmd<-paste("R --no-save --args ",longlat[1]," ",longlat[2]," < extract.R",sep='')
system(cmd)
soilpro<-read.csv('data.csv')
setwd(prevdir)
soilpro[,1]<-c(2.5,7.5,22.5,45,80,150)
colnames(soilpro)[1] <- 'depth'
```

Now get hydraulic properties for this soil using Cosby et al. 1984 pedotransfer functinos.

```{r, eval=FALSE}
DEP <- c(0., 1.,  3, 5.,  10,  15,  30.,  60.,  100.,  200.) # Soil nodes (cm)
soil.hydro<-pedotransfer(soilpro = as.data.frame(soilpro), DEP = DEP)
PE<-soil.hydro$PE
BB<-soil.hydro$BB
BD<-soil.hydro$BD
KS<-soil.hydro$KS
BD[1:9]<-BD[1:9]#*1.7 # make top 15 cm more dense
BulkDensity <- BD[seq(1,19,2)]*1000 #soil bulk density, kg/m3
```

Get qualitative soil type out of interest.

```{r, eval=FALSE}
# search through observed textures and find the nearest match to Campell and Norman's Table 9.1
stypes<-NULL
for(m in 1:nrow(soilpro)){
 ssq<-(CampNormTbl9_1[,2]-soilpro[m,4]/100)^2 + (CampNormTbl9_1[,3]-soilpro[m,3]/100)^2
 stypes[m]<-which(ssq==min(ssq))
}

# produce a table of the qualitative soil profile
soils<-as.character(CampNormTbl9_1[,1])
profile<-as.data.frame(cbind(soilpro[,1],soils[stypes]), stringsAsFactors=FALSE)
profile[,1]<-as.numeric(profile[,1])
colnames(profile)<-c("depth","soiltype")
library(knitr)
kable(profile)
``` 

Get LAI
```{r, eval=FALSE}
library(raster)
monthly_LAI <- stack("W:/monthlyLAI.nc")
lonlat<-as.data.frame(longlat)
monthly_LAI <- crop(monthly_LAI, extent(lonlat[1,1]-0.1, lonlat[1,1]+0.1, lonlat[1,2]-0.1, lonlat[1,2]+0.1))
LAI_month<-extract(monthly_LAI, cbind(longlat[1],longlat[2]))
juldays12 <- c(15, 46, 74, 105, 135, 166, 196, 227, 258, 288, 319, 349)
LAI <- suppressWarnings(spline(juldays12, LAI_month, n = 365, xmin = 1, xmax = 365, method = "periodic"))$y
LAI <- rep(LAI, nyears)
```

Now run the microclimate model and get the output.

```{r, eval=FALSE}
micro<-micro_aust(loc = loc, maxshade = 70, ystart = ystart, yfinish = yfinish, PE = PE, BB = BB, BD = BD, KS = KS, BulkDensity = BulkDensity, spatial = "w:/", windfac = 1, LAI = LAI, cap = 0, TIMAXS = c(3, 1, 0, 0))
save(micro, file = "micro.Rda")
```

Estimate burrow humidity as average vapour pressure between near-surface and soil

```{r}
load("micro.Rda")

sat<-WETAIR(db = micro$soil[,3:12])$esat # saturated soil vapour pressure
vpres_soil<-sat*micro$humid[,3:12]
e<-WETAIR(db = micro$metout[,3], rh = micro$metout[,5])$e # ambient vapour pressure
satmean <- vpres_soil
weights <- seq(0, .9,.1)
satmean <- t(apply(cbind(e,e,e,e,e,e,e,e,e,e), 1, function(x)x*(1-weights))) + t(apply(vpres_soil, 1, function(x)x*weights))
#satmean <- (e * 0.5 + vpres_soil * 0.5)  # mean of ambient and saturated vapour pressure
newRH<-satmean/sat
newRH[newRH>1]<-1
micro$humid[,3:12]<-newRH#micro$metout[,5]/100 # assume ambient humidity down to 30cm
sat<-WETAIR(db = micro$shadsoil[,3:12])$esat # saturated soil vapour pressure
e<-WETAIR(db = micro$shadmet[,3], rh = micro$shadmet[,5])$e # ambient vapour pressure
satmean <- vpres_soil
satmean <- t(apply(cbind(e,e,e,e,e,e,e,e,e,e), 1, function(x)x*(1-weights))) + t(apply(vpres_soil, 1, function(x)x*weights))#satmean <- (e * 0.5 + vpres_soil * 0.5)  # mean of ambient and saturated vapour pressure
newRH<-satmean/sat
newRH[newRH>1]<-1
micro$shadhumid[,3:12]<-newRH#micro$metout[,5]/100 # assume ambient humidity down to 30cm
```

Next, append dates

```{r}
tzone=paste("Etc/GMT-",10,sep="") # doing it this way ignores daylight savings!
dates=seq(ISOdate(ystart,1,1,tz=tzone)-3600*12, ISOdate((ystart+nyears),1,1,tz=tzone)-3600*13, by="hours")
dates=subset(dates, format(dates, "%m/%d")!= "02/29") # remove leap years
dates2=seq(ISOdate(ystart,1,1,tz=tzone)-3600*12, ISOdate((ystart+nyears),1,1,tz=tzone)-3600*13, by="days") 
dates2=subset(dates2, format(dates2, "%m/%d")!= "02/29") # remove leap years

metout <- as.data.frame(cbind(dates,as.data.frame(micro$metout))) # above ground microclimatic conditions, min shade
shadmet <- as.data.frame(cbind(dates,as.data.frame(micro$shadmet)))
soil <- as.data.frame(cbind(dates,as.data.frame(micro$soil))) # soil temperatures, minimum shade
shadsoil <- as.data.frame(cbind(dates,as.data.frame(micro$shadsoil))) # soil temperatures, maximum shade
humid <- as.data.frame(cbind(dates,as.data.frame(micro$humid))) # soil temperatures, minimum shade
shadhumid <- as.data.frame(cbind(dates,as.data.frame(micro$shadhumid))) # soil temperatures, maximum shade
```


##compare with measurements

```{r}

############################# read in weather data ###################################################
weather_obs = read.csv("csv summaries/weather_obs.csv", stringsAsFactors = FALSE)[,-1]
weather_obs$TIMESTAMP = as.POSIXct(weather_obs$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S")

############################# read in copper model data and format dates ###################################################

# read in data
obs.surfaces=read.csv("csv summaries/all.soil.surfaces.csv",stringsAsFactors = FALSE)[,-1]
obs.5cms=read.csv("csv summaries/all.soil.5cms.csv",stringsAsFactors = FALSE)[,-1]
obs.15cms=read.csv("csv summaries/all.soil.15cms.csv",stringsAsFactors = FALSE)[,-1]
obs.30cms=read.csv("csv summaries/all.soil.30cms.csv",stringsAsFactors = FALSE)[,-1]
obs.50cms=read.csv("csv summaries/all.soil.50cms.csv",stringsAsFactors = FALSE)[,-1]
obs.1ms=read.csv("csv summaries/all.soil.1ms.csv",stringsAsFactors = FALSE)[,-1]

obs.burrow.surf=read.csv("csv summaries/all.burrow.surf.csv",stringsAsFactors = FALSE)[,-1]
obs.burrow.mid=read.csv("csv summaries/all.burrow.mid.csv",stringsAsFactors = FALSE)[,-1]
obs.burrow.deep=read.csv("csv summaries/all.burrow.deep.csv",stringsAsFactors = FALSE)[,-1]

source('addTrans.R') # function for making transparent plots
source('polylims.R') # function for creating 
# read in data
mean.max.surf=read.csv("csv summaries/mean.max.surf.csv",stringsAsFactors = FALSE)[,-1]
mean.max.mid=read.csv("csv summaries/mean.max.mid.csv",stringsAsFactors = FALSE)[,-1]
mean.max.deep=read.csv("csv summaries/mean.max.deep.csv",stringsAsFactors = FALSE)[,-1]

mean.min.surf=read.csv("csv summaries/mean.min.surf.csv",stringsAsFactors = FALSE)[,-1]
mean.min.mid=read.csv("csv summaries/mean.min.mid.csv",stringsAsFactors = FALSE)[,-1]
mean.min.deep=read.csv("csv summaries/mean.min.deep.csv",stringsAsFactors = FALSE)[,-1]

# format dates
mean.max.surf$date = as.POSIXct(mean.max.surf$date, format = "%Y-%m-%d")
mean.max.mid$date = as.POSIXct(mean.max.mid$date, format = "%Y-%m-%d")
mean.max.deep$date = as.POSIXct(mean.max.deep$date, format = "%Y-%m-%d")

mean.min.surf$date = as.POSIXct(mean.min.surf$date, format = "%Y-%m-%d")
mean.min.mid$date = as.POSIXct(mean.min.mid$date, format = "%Y-%m-%d")
mean.min.deep$date = as.POSIXct(mean.min.deep$date, format = "%Y-%m-%d")

# create polygons for weather surf humidity
ts <- seq.POSIXt(as.POSIXlt(min(mean.max.surf$date)), as.POSIXlt(max(mean.max.surf$date)), by="day")
ts <- format.POSIXct(ts,'%Y-%m-%d')
mean.max.surf$merge <- format.POSIXct(mean.max.surf$date,'%Y-%m-%d')
df <- data.frame(merge=ts)
mean.max.surf2 <- full_join(df,mean.max.surf)
mean.max.surf2$merge <- as.POSIXct(mean.max.surf2$merge, format = "%Y-%m-%d")

ts <- seq.POSIXt(as.POSIXlt(min(mean.min.surf$date)), as.POSIXlt(max(mean.min.surf$date)), by="day")
ts <- format.POSIXct(ts,'%Y-%m-%d')
mean.min.surf$merge <- format.POSIXct(mean.min.surf$date,'%Y-%m-%d')
df <- data.frame(merge=ts)
mean.min.surf2 <- full_join(df,mean.min.surf)
mean.min.surf2$merge <- as.POSIXct(mean.min.surf2$merge, format = "%Y-%m-%d")

# create polygons for weather mid humidity
ts <- seq.POSIXt(as.POSIXlt(min(mean.max.mid$date)), as.POSIXlt(max(mean.max.mid$date)), by="day")
ts <- format.POSIXct(ts,'%Y-%m-%d')
mean.max.mid$merge <- format.POSIXct(mean.max.mid$date,'%Y-%m-%d')
df <- data.frame(merge=ts)
mean.max.mid2 <- full_join(df,mean.max.mid)
mean.max.mid2$merge <- as.POSIXct(mean.max.mid2$merge, format = "%Y-%m-%d")

ts <- seq.POSIXt(as.POSIXlt(min(mean.min.mid$date)), as.POSIXlt(max(mean.min.mid$date)), by="day")
ts <- format.POSIXct(ts,'%Y-%m-%d')
mean.min.mid$merge <- format.POSIXct(mean.min.mid$date,'%Y-%m-%d')
df <- data.frame(merge=ts)
mean.min.mid2 <- full_join(df,mean.min.mid)
mean.min.mid2$merge <- as.POSIXct(mean.min.mid2$merge, format = "%Y-%m-%d")

# create polygons for weather deep humidity
ts <- seq.POSIXt(as.POSIXlt(min(mean.max.deep$date)), as.POSIXlt(max(mean.max.deep$date)), by="day")
ts <- format.POSIXct(ts,'%Y-%m-%d')
mean.max.deep$merge <- format.POSIXct(mean.max.deep$date,'%Y-%m-%d')
df <- data.frame(merge=ts)
mean.max.deep2 <- full_join(df,mean.max.deep)
mean.max.deep2$merge <- as.POSIXct(mean.max.deep2$merge, format = "%Y-%m-%d")

ts <- seq.POSIXt(as.POSIXlt(min(mean.min.deep$date)), as.POSIXlt(max(mean.min.deep$date)), by="day")
ts <- format.POSIXct(ts,'%Y-%m-%d')
mean.min.deep$merge <- format.POSIXct(mean.min.deep$date,'%Y-%m-%d')
df <- data.frame(merge=ts)
mean.min.deep2 <- full_join(df,mean.min.deep)
mean.min.deep2$merge <- as.POSIXct(mean.min.deep2$merge, format = "%Y-%m-%d")

xlims <- c(as.POSIXct("2013-10-01 00:00:00", format = "%Y-%m-%d %H:%M:%S"), as.POSIXct("2014-09-30 00:00:00", format = "%Y-%m-%d %H:%M:%S"))

par(mfrow = c(2,1))
par(mar = c(1,1,1,1) + 0.1) # margin spacing stuff 
par(mar = c(4,4,1,4) + 0.1) # margin spacing stuff
plot(metout$dates, metout$TAREF, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(1.2), ' m', ' air temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(weather_obs$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S"), weather_obs$AirTemp_C_Avg,col=addTrans("red",150))

plot(metout$dates, metout$VREF, type = 'l', xlim = xlims, ylim = c(0,10), ylab = "1.2 m wind speed (m/s)", xlab ='date')
lines(as.POSIXct(weather_obs$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S"), weather_obs$WindSpeed_ms_Avg,col=addTrans("red",150))

plot(metout$dates, metout$RH, type = 'l', xlim = xlims, ylim = c(0,100), ylab = "1.2 m relative humidity (%)", xlab ='date')
lines(as.POSIXct(weather_obs$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S"), weather_obs$RH_Avg,col=addTrans("red",150))

plot(metout$dates, metout$SOLR, type = 'l', xlim = xlims, ylim = c(0,1300), ylab = "solar radiation (w / m2)", xlab ='date')
lines(as.POSIXct(weather_obs$TIMESTAMP, format = "%Y-%m-%d %H:%M:%S"), weather_obs$Solar_Avg,col=addTrans("red",150))

par(mfrow = c(3,1))
par(mar = c(1,1,1,1) + 0.1) # margin spacing stuff 
par(mar = c(4,4,1,4) + 0.1) # margin spacing stuff

plot(soil$dates, soil$D0cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(0), ' cm', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(obs.surfaces$date_time, format = "%Y-%m-%d %H:%M:%S"), obs.surfaces$temperature,col=addTrans("red",150))

plot(soil$dates, soil$D5cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(5), ' cm', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(obs.5cms$date_time, format = "%Y-%m-%d %H:%M:%S"), obs.5cms$temperature,col=addTrans("red",150))

plot(soil$dates, soil$D15cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(15), ' cm', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(obs.15cms$date_time, format = "%Y-%m-%d %H:%M:%S"), obs.15cms$temperature,col=addTrans("red",150))

plot(soil$dates, soil$D30cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(5), ' cm', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(obs.30cms$date_time, format = "%Y-%m-%d %H:%M:%S"), obs.30cms$temperature,col=addTrans("red",150))

plot(soil$dates, soil$D50cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(50), ' cm', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(obs.50cms$date_time, format = "%Y-%m-%d %H:%M:%S"), obs.50cms$temperature,col=addTrans("red",150))

plot(soil$dates, soil$D100cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(1), ' m', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(obs.1ms$date_time, format = "%Y-%m-%d %H:%M:%S"), obs.1ms$temperature,col=addTrans("red",150))

par(mfrow = c(3,1))
par(mar = c(1,1,1,1) + 0.1) # margin spacing stuff 
par(mar = c(4,4,1,4) + 0.1) # margin spacing stuff

sitestodo <- unique(obs.burrow.deep$site)

for(i in 1:length(sitestodo)){
  
plot.obs.burrow.surf <- subset(obs.burrow.surf, site == sitestodo[i])
plot.obs.burrow.mid <- subset(obs.burrow.mid, site == sitestodo[i])
plot.obs.burrow.deep <- subset(obs.burrow.deep, site == sitestodo[i])

plot(soil$dates, soil$D5cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(5), ' cm', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(plot.obs.burrow.surf$date_time, format = "%Y-%m-%d %H:%M:%S"), plot.obs.burrow.surf$temperature,col=addTrans("red",150))

plot(soil$dates, soil$D30cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(30), ' cm', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(plot.obs.burrow.mid$date_time, format = "%Y-%m-%d %H:%M:%S"), plot.obs.burrow.mid$temperature,col=addTrans("red",150))

plot(soil$dates, soil$D100cm, type = 'l', xlim = xlims, ylim = c(-10,70), ylab = bquote(paste(.(1), ' m', ' temperature (', ~degree~C,')')), xlab ='date')
lines(as.POSIXct(plot.obs.burrow.deep$date_time, format = "%Y-%m-%d %H:%M:%S"), plot.obs.burrow.deep$temperature,col=addTrans("red",150))
}

par(mfrow = c(3,1))
par(mar = c(1,1,1,1) + 0.1) # margin spacing stuff 
par(mar = c(4,4,1,4) + 0.1) # margin spacing stuff


plot(humid$dates, humid$RH5cm*100, type = 'l', xlim = xlims, ylim = c(0,100), ylab = "burrow humidity shallow, %", xlab ='date')
# add the polygons
polys = polylims(mean.max.surf2$merge, mean.min.surf2$mean.min.humidity, mean.max.surf2$mean.max.humidity)
for(i in 1:length(polys)){
	polygon(polys[[i]]$x, polys[[i]]$y,col=addTrans("red",150), border = NA)
}

plot(humid$dates, humid$RH30cm*100, type = 'l', xlim = xlims, ylim = c(0,100), ylab = "burrow humidity mid, %", xlab ='date')
# add the polygons
polys = polylims(mean.max.mid2$merge, mean.min.mid2$mean.min.humidity, mean.max.mid2$mean.max.humidity)
for(i in 1:length(polys)){
	polygon(polys[[i]]$x, polys[[i]]$y, col=addTrans("red",150), border = NA)
}

plot(humid$dates, humid$RH100cm*100, type = 'l', xlim = xlims, ylim = c(0,100), ylab = "burrow humidity deep, %", xlab ='date')
# add the polygons
polys = polylims(mean.max.deep2$merge, mean.min.deep2$mean.min.humidity, mean.max.deep2$mean.max.humidity)
for(i in 1:length(polys)){
	polygon(polys[[i]]$x, polys[[i]]$y,col=addTrans("red",150), border = NA)
}


sitestodo <- unique(obs.burrow.deep$site)

for(i in 1:length(sitestodo)){
  
plot.obs.burrow.surf <- subset(obs.burrow.surf, site == sitestodo[i])
plot.obs.burrow.mid <- subset(obs.burrow.mid, site == sitestodo[i])
plot.obs.burrow.deep <- subset(obs.burrow.deep, site == sitestodo[i])

plot(humid$dates, humid$RH5cm*100, type = 'l', xlim = xlims, ylim = c(0,100), ylab = "burrow humidity shallow, %", xlab ='date')
lines(as.POSIXct(plot.obs.burrow.surf$date_time, format = "%Y-%m-%d %H:%M:%S"), plot.obs.burrow.surf$humidity,col=addTrans("red",150))


plot(humid$dates, humid$RH30cm*100, type = 'l', xlim = xlims, ylim = c(0,100), ylab = "burrow humidity mid, %", xlab ='date')
lines(as.POSIXct(plot.obs.burrow.mid$date_time, format = "%Y-%m-%d %H:%M:%S"), plot.obs.burrow.mid$humidity,col=addTrans("red",150))


plot(humid$dates, humid$RH100cm*100, type = 'l', xlim = xlims, ylim = c(0,100), ylab = "burrow humidity deep, %", xlab ='date')
lines(as.POSIXct(plot.obs.burrow.deep$date_time, format = "%Y-%m-%d %H:%M:%S"), plot.obs.burrow.deep$humidity,col=addTrans("red",150))

}
```
